---
title: "PlanetDataNDVIStuff"
format: html
editor: visual
---

## Introduction

I am going to try to get some NDVI from the planet data

## Load the Data

```{r LoadPackages}
#| message: false
#| warning: false

library(tidyverse)
library(terra)
library(purrr)
```

```{r LoadData}

# Loads in some planet data
PlanetData1 <- rast("./data/PlanetData/PSScene/20240620_182805_39_2473_3B_AnalyticMS_SR_clip.tif")
PlanetData2 <- rast("./data/PlanetData/PSScene/20240620_182807_25_2473_3B_AnalyticMS_SR_clip.tif") |> 
  project("epsg:4326")

# Scale it back down
PlanetData2 <- PlanetData2 * 0.0001

# Loads in the fields
ReinkeWheelLine <- vect("./data/FieldData/onx-markups-03132026.kml")
LittleReinkePivot <- vect("./data/FieldData/onx-markups-03132026(1).kml")
NReinkeWheelLine <- vect("./data/FieldData/onx-markups-03132026(2).kml")
WilsonPivot <- vect("./data/FieldData/onx-markups-03132026(3).kml")
NorthReinkePivot <- vect("./data/FieldData/onx-markups-03132026(4).kml")
SouthReinkePivot <- vect("./data/FieldData/onx-markups-03132026(5).kml")
UpperPivot <- vect("./data/FieldData/onx-markups-03132026(6).kml")
ClarePivot <- vect("./data/FieldData/onx-markups-03132026(7).kml")
LittleValleyPivot <- vect("./data/FieldData/onx-markups-03132026(8).kml")
SnubPivot <- vect("./data/FieldData/onx-markups-03132026(9).kml")
TurkeyBarnPivot <- vect("./data/FieldData/onx-markups-03132026(10).kml")
BigValleyPivot <- vect("./data/FieldData/onx-markups-03132026(11).kml")
NewValleyPivot <- vect("./data/FieldData/onx-markups-03132026(12).kml")
HousePivot <- vect("./data/FieldData/onx-markups-03132026(13).kml")

# Combines the fields into one vector
combined <- rbind(ReinkeWheelLine, LittleReinkePivot, NReinkeWheelLine, WilsonPivot, NorthReinkePivot, SouthReinkePivot, UpperPivot, ClarePivot, LittleValleyPivot, SnubPivot, TurkeyBarnPivot, BigValleyPivot, NewValleyPivot, HousePivot)

# Plots the fields with their names
plot(combined)
text(combined, 'Name', cex = 0.5 )
```

That's all well and good, but let's see if I can auto-load all the files I want.

```{r Automation}

# List the files in the chosen directory that contain the regex I am looking for
fileList <- list.files(path = "./data/PlanetData/PSScene/", pattern = "SR", full.names = T)

#That is a list of files. Now to generate rasters from them
rastList <- map(fileList, rast)

# Now let's check one
plot(rastList[[3]])
```

Great success (*Borat voice*)

```{r MosaicRasters}

# Now we're going to mosaic some rasters
juneRast <- mosaic(rastList[[1]], rastList[[2]], rastList[[3]], rastList[[4]], rastList[[5]])

julyRast <- mosaic(rastList[[6]], rastList[[7]])

augRast <- mosaic(rastList[[8]], rastList[[9]], rastList[[10]], rastList[[11]])

septRast <- mosaic(rastList[[12]], rastList[[13]], rastList[[14]], rastList[[15]], rastList[[16]])

octRast <- mosaic(rastList[[17]], rastList[[18]], rastList[[19]], rastList[[20]], rastList[[21]])

novRast <- mosaic(rastList[[22]], rastList[[23]], rastList[[24]], rastList[[25]], rastList[[26]], rastList[[27]], rastList[[28]], rastList[[29]])

# And a check
plotRGB(augRast, r = 3, b = 1, g = 2, stretch = "lin")
```

Gorgeous. It works. I now have half the year's rasters.

Okay, that takes care of getting some data in. Now lets plot them in over each other.

```{r Overlay}

plotRGB(PlanetData2, r = 3, b = 1, g = 2, stretch = "lin")
polys(combined, col = "tomato", lwd = 1, alpha = 0.5)
```

Now, let's do an NDVI.

```{r NDVI}

# Raster math
NDVI <- (PlanetData2$nir - PlanetData2$red) / (PlanetData2$nir + PlanetData2$red)
plot(NDVI, col = map.pal("grass", 100))
```

Now, let's clip a field out and hit it with some G\*.

```{r Clip}

# Clip out one field
BigValleyNDVI <- crop(NDVI, BigValleyPivot, mask = T)
ClareNDVI <- crop(NDVI, ClarePivot, mask = T)
HouseNDVI <- crop(NDVI, HousePivot, mask = T)
LittleReinkeNDVI <- crop(NDVI, LittleReinkePivot, mask = T)
LittleValleyNDVI <- crop(NDVI, LittleValleyPivot, mask = T)
NewValleyNDVI <- crop(NDVI, NewValleyPivot, mask = T)
NorthReinkeNDVI <- crop(NDVI, NorthReinkePivot, mask = T)
NReinkeWheelNDVI <- crop(NDVI, NReinkeWheelLine, mask = T)
ReinkeWheelNDVI <- crop(NDVI, ReinkeWheelLine, mask = T)
SnubNDVI <- crop(NDVI, SnubPivot, mask = T)
SouthReinkeNDVI <- crop(NDVI, SouthReinkePivot, mask = T)
TurkeyBarnNDVI <- crop(NDVI, TurkeyBarnPivot, mask = T)
UpperNDVI <- crop(NDVI, UpperPivot, mask = T)
WilsonNDVI <- crop(NDVI, WilsonPivot, mask = T)

NDVIlist <- c(BigValleyNDVI, ClareNDVI, HouseNDVI, LittleValleyNDVI, LittleReinkeNDVI, NewValleyNDVI, NorthReinkeNDVI, NReinkeWheelNDVI, ReinkeWheelNDVI, SnubNDVI, TurkeyBarnNDVI, UpperNDVI, WilsonNDVI)

plot(BigValleyNDVI)
```

And now G\*.

```{r LocalMoran}

# Run GStar
LocMoran <- autocor(BigValleyNDVI, method = "moran", global = F)

# Set up a map function
map(NDVIlist, autocor())

plot(GStar)
```
