---
title: "Sentinel_Soil_Moisture"
format: html
editor: visual
---

## Sourcing Data

Based on my preliminary readings, it seems like the package `rstac` is going to be my best source for pre-processed and analysis-ready data.

```{r IntroToRSTAC}
#| message: false
#| warning: false

library(rstac)

# Sets the STAC to Microsoft's Planetary Computer
stac_source <- stac(
  "https://planetarycomputer.microsoft.com/api/stac/v1"
)

# This one will list the data collections available in that STAC
collections_query <- stac_source |> 
  collections()

# This runs the collection query and saves the result
coll_list <- get_request(collections_query)

library(tidyverse)
library(purrr)

readable_list <- tibble(
  collection = map_chr(coll_list$collections, "id")
)
```

Now, I need a bounding box. To make that bounding box, I should probably pull in the onX point data of the fields and use the extent of that to create the bounding box for my rstac data query. Then, I will need to convert the points to lines and polygons in order to use the field polygons to clip the sentinel data to just the fields that I want to look at.

```{r LoadData}
#| message: false
#| warning: false

library(terra)

# Loads in all the field data
ReinkeWheelLine <- vect("./data/FieldData/onx-markups-03132026.kml")
LittleReinkePivot <- vect("./data/FieldData/onx-markups-03132026(1).kml")
NReinkeWheelLine <- vect("./data/FieldData/onx-markups-03132026(2).kml")
WilsonPivot <- vect("./data/FieldData/onx-markups-03132026(3).kml")
NorthReinkePivot <- vect("./data/FieldData/onx-markups-03132026(4).kml")
SouthReinkePivot <- vect("./data/FieldData/onx-markups-03132026(5).kml")
UpperPivot <- vect("./data/FieldData/onx-markups-03132026(6).kml")
ClarePivot <- vect("./data/FieldData/onx-markups-03132026(7).kml")
LittleValleyPivot <- vect("./data/FieldData/onx-markups-03132026(8).kml")
SnubPivot <- vect("./data/FieldData/onx-markups-03132026(9).kml")
TurkeyBarnPivot <- vect("./data/FieldData/onx-markups-03132026(10).kml")
BigValleyPivot <- vect("./data/FieldData/onx-markups-03132026(11).kml")
NewValleyPivot <- vect("./data/FieldData/onx-markups-03132026(12).kml")
HousePivot <- vect("./data/FieldData/onx-markups-03132026(13).kml")

# SHOW ME THE MONEY
plot(c(ReinkeWheelLine, LittleReinkePivot, NReinkeWheelLine, WilsonPivot, NorthReinkePivot, SouthReinkePivot, UpperPivot, ClarePivot, LittleValleyPivot, SnubPivot, TurkeyBarnPivot, BigValleyPivot, NewValleyPivot, HousePivot))
```

Well, that didn't quite work like I'd hoped, but it did work well enough to see that there are some problems with this data. Namely, the polygons aren't all clean and round like they should be. Not an immediate problem, but I may need to fix them with some satellite-based geo-referencing before this is over.

```{r CombineData}

# Combines the fields into one spatVector
combined <- rbind(ReinkeWheelLine, LittleReinkePivot, NReinkeWheelLine, WilsonPivot, NorthReinkePivot, SouthReinkePivot, UpperPivot, ClarePivot, LittleValleyPivot, SnubPivot, TurkeyBarnPivot, BigValleyPivot, NewValleyPivot, HousePivot)

# Gets the extent of that spatVector
FieldExt <- ext(combined)

# SHOW ME THE MONEY
plot(combined)
text(combined, 'Name', cex = 0.5 )
```

Okay, well, that's purdy. Got the extent I'm working with. Should be able to go back to `rstac` now and pull some sentinel-1 data.

```{r PullRSTAC}

# Structures the STAC query
stac_query <- stac_search(
  q = stac_source,
  collection = "sentinel-1-rtc",
  bbox = FieldExt[c(1, 3, 2, 4)],
  datetime = "2025-04-01/2025-04-05"
)

# Runs the STAC query
stac_list <- get_request(stac_query)

# Signs the STAC query so Planetary Computer will let me download items
signed_stac_query <- items_sign(
  stac_list,
  sign_planetary_computer()
)

# Downloads a rendered preview
assets_download(signed_stac_query, c("vh", "vv"), 
                output_dir = "./data/RadarData/", overwrite = T)
```

Before we go on, it's worth taking a moment to pause and think about data structures and goals here. I want to download just the data that covers my fields, so I want to use the GDAL system described on the [STAC website](https://stacspec.org/en/tutorials/1-download-data-using-r/) tutorial for using `rstac`. We also need to do some reading about the naming conventions for sentinel data (S1A vs. S1C), and what assets we want (VH, VV). So let's do that and regroup in five.

Okay, let's have a discussion of the strengths and limits of Sentinel-based soil moisture analysis. It is definitely possible to do with a meaningful degree of accuracy, but it requires a few things. First, it requires LAI (Leaf Area Index) control points, which is most accurately determined by doing some insane vegetation cutting and measuring. That's a lot, so I am hoping that since the fields are hay fields with homogeneous grass cover we can get away with estimating LAI from remote sensing data.

Second, there needs to be some input for soil moisture to make this data meaningful. This basically just means soil moisture sensors, which are relatively cheap and easy to install, particularly in hay fields that aren't being tilled every year.

#### (Sound of Brakes Screeching to a Halt)

Scratch all that. Soil moisture may not be possible. Let's just play with the radar data.

```{r RadExp}

# Load in the VH raster
radRastVH <- rast("./data/RadarData/sentinel1-grd-rtc/GRD/2025/4/1/IW/DV/S1C_IW_GRDH_1SDV_20250401T011917_20250401T011942_001693_002D8A_88A4/measurement/iw-vh.rtc.tiff") |> 
  project("epsg:4326")

# Load in the VV raster
radRastVV <- rast("./data/RadarData/sentinel1-grd-rtc/GRD/2025/4/1/IW/DV/S1C_IW_GRDH_1SDV_20250401T011917_20250401T011942_001693_002D8A_88A4/measurement/iw-vv.rtc.tiff") |> 
  project("epsg:4326")

# Plot the raster with the fields
plot(radRastVH)
polys(combined)

# Cropping the rasters to the fields
radRastVHcrop <- crop(radRastVH, combined)
radRastVVcrop <- crop(radRastVV, combined)

# Looking at the cropped raster
plot(radRastVHcrop, col = map.pal("haxby"))
polys(combined)

# Looking at the Big Valley pivot
bigValleyCrop <- crop(radRastVVcrop, BigValleyPivot, mask = T)
plot(bigValleyCrop)

#Looking at the Snub pivot
snubCrop <- crop(radRastVVcrop, SnubPivot, mask = T)
plot(snubCrop)

# Looking at all the pivots
combinedCrop <- crop(radRastVV, combined, mask = T)
plot(combinedCrop)
```

Okay, i confirmed with landsat data that the difference we can see in the Big Valley pivot is indeed because the irrigation system is running on one side, but not the other. I know this because the soil is darker on the irrigated side. I want to do a time series here, but I will need to improve the STAC query because the files I downloaded with the other query were far too large. I want to only download the portion that overlaps with a bounding box drawn around my fields.

```{r ImprovedSTAC}

# Sets up a stac query to pull all of April 2024
improvedSTACquery <- stac(
  "https://planetarycomputer.microsoft.com/api/stac/v1"
) |>
  stac_search(
    collections = "sentinel-1-rtc",
    bbox = FieldExt[c(1, 3, 2, 4)],
    datetime = "2024-01-01/2024-12-31"
  ) |>
  get_request()

# A function that sets up the GDAL request
make_vsicurl_url <- function(base_url) {
  paste0(
    "/vsicurl", 
    "?pc_url_signing=yes",
    "&pc_collection=sentinel-1-rtc",
    "&url=",
    base_url
  )
}

# Makes the URLs, identifies the relative orbit, drops the ones I don't want, then removes the NAs from the vector
sentinel_urls <- ifelse(
  relOrbit(as.integer(str_extract(
    make_vsicurl_url(assets_url(improvedSTACquery, "vv")),
    "(?<=_)[:digit:]{6}(?=_)"))) == 100,
  make_vsicurl_url(assets_url(improvedSTACquery, "vv")),
  NA) |> 
  stats::na.omit()
  


# Downloads just the bit of data I want. Much faster.
map(sentinel_urls, ~sf::gdal_utils(
  "warp",
  source = .x,
  destination = paste0("C:/Users/Luke/OneDrive/Documents/School/Spring_2026/GPHY_484/Project/Sentinel_Soil_Moisture/data/RadarData/TimeSeriesApril/", 
                       str_extract(.x, "[:digit:]{8}"), ".tif"),
  options = c(
    "-t_srs", crs(combined),
    "-te", FieldExt[c(1, 3, 2, 4)]
  )
))


```

```{r Test}

radRastApr10 <- rast("./data/RadarData/TimeSeriesApril/20240410.tif")

crop(radRastApr10, combined, mask = T) |> 
  plot()
```

```{r RelativeOrbit}

# Defines a function that computes relative orbit from absolute orbit based on the formula given by the ESA
relOrbit <- function(absOrbit) {
  ((absOrbit - 73) %% 175) + 1
}

# Checks the relative orbits of my sentinel urls to make sure they're all the ones I want
map(sentinel_urls, ~relOrbit(as.integer(str_extract(.x, "(?<=_)[:digit:]{6}(?=_)"))))
```

Well, now that I have the data, I can tell that I can for sure see the fields in the data, and I think I can see some correlation with irrigation on the April Big Valley Pivot images, because at that time the LAI is basically zero but based on some visual inspections of concurrent landsat imagery, I can say that they had already begun irrigation. Therefore, the only thing that is different between the two halves of the Big Valley Pivot is that one half is wet and the other half is dry. So we might be back! Even if translating this into some actual number may be challenging, I do believe that this convinces me that radar is still worthwhile to pursue.

Now I need to pull all these files in as rasters, put them in a list, and all that jazz.

```{r LoadingRasters}

# Pulls the file paths
fileList <- list.files("./data/RadarData/TimeSeriesApril/", full.names = T)

# Pulls names out of the files for later
namesList <- str_extract(fileList, pattern = "[:digit:]{8}")

# Generates the raster list
rastList <- map(fileList, ~rast(.x))

# Names each element
rastList <- setNames(rastList, namesList)

# Test
plot(rastList[[1]])
```

Okay, the rasters are loaded in and set up as named elements in a list. The projection is correct, and the rasters appear to be showing up properly. Now I need to crop out each field for each raster, which should give me 392 individual rasters. To do that, I will need a list of fields as well, then I should be able to do this with nested map() functions.

```{r Cropping}

# list field files
fieldFileList <- list.files(path = "./data/FieldData/", full.names = T)

# Turn them into spatVectors
vectList <- map(fieldFileList, vect)

# Extract field names
fieldNames <- map_chr(vectList, ~.x[]$Name)

# Assigns the fields their names
vectList <- setNames(vectList, fieldNames)

# Writing a crop function
cropFunc <- function(raster) {
  map(vectList, ~crop(raster, .x, mask = T))
}

# Doing the crop
croppedFields <- map(rastList, ~cropFunc(.x))

# Test
plot(croppedFields[[13]][[3]])
```

Cool. I'd say that works. I will also say, I don't have a good explanation for why there is a split in the Big Valley Pivot's radar backscatter in January and February. That will take further explanation and would probably be aided by some optical spectrum images of the field at that time. Anyways, lets build the ggplot2 animation and see what that looks like.

```{r asDataframe}
library(ggh4x)
library(gganimate)

# Function to convert raster to a data frame with columns for month and field.
dfFunc <- function(rasterlist, date) {
  imap(rasterlist, \(x, field) {
    as.data.frame(x, xy = T) |> 
      mutate(
        date = date,
        field = field
      )
  })
}

# Create the list of data frames
bigData <- imap(croppedFields, dfFunc) |> 
  map(~list_rbind(.x)) |> 
  list_rbind() |> 
  mutate(
    date = ymd(date)
  ) |> 
  rename(
    value = `Sentinel-1 Calibrated and Terrain Corrected VV`
  )

# Set the design
design <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, NA, 13, 14, NA), 
                 4, 4, byrow = T)

# Test the plot
bigData |> 
  filter(date == "2024-07-15") |> 
  ggplot() +
    geom_tile(aes(x = x, y = y, fill = value)) +
    scale_fill_viridis_c(option = "A") +
    facet_manual(~field, design = design, scales = "free") +
    labs(
      title = "AMB West Radar Backscatter",
      subtitle = "Date: {frame_time}",
      x = "",
      y = "",
      caption = "From Sentinel-1 VV 30m data, 01/24-12/24\nBy Luke Petrus."
    ) +
    theme_bw() +
    theme(axis.text = element_blank(),
          plot.title = element_text(face = "bold", hjust = 0.5, color = "black"),
          plot.subtitle = element_text(hjust = 0.5, color = "black"),
          plot.background = element_rect(fill = "white"),
          legend.background = element_rect(fill = "white"),
          legend.text = element_text(color = "black"),
          legend.title = element_text(color = "black"),
          plot.caption = element_text(hjust = 0)
    )

# Now for the animation
animPlot <-   ggplot() +
  geom_tile(aes(x = x, y = y, fill = value)) +
  facet_manual(~field, design = design, scales = "free") +
  labs(
    title = "AMB West Radar Backscatter",
    subtitle = "Date: {frame_time}",
    x = "",
    y = ""
  ) +
  theme_bw() +
  theme(axis.text = element_blank(),
        plot.title = element_text(face = "bold", hjust = 0.5, color = "black"),
        plot.subtitle = element_text(hjust = 0.5, color = "black"),
        plot.background = element_rect(fill = "white"),
        legend.background = element_rect(fill = "white"),
        legend.text = element_text(color = "black"),
        legend.title = element_text(color = "black")
  ) +
  transition_time(date)
```
