---
title: "Sentinel_Soil_Moisture"
format: html
editor: visual
---

## Sourcing Data

Based on my preliminary readings, it seems like the package `rstac` is going to be my best source for pre-processed and analysis-ready data.

```{r IntroToRSTAC}
#| message: false
#| warning: false

library(rstac)

# Sets the STAC to Microsoft's Planetary Computer
stac_source <- stac(
  "https://planetarycomputer.microsoft.com/api/stac/v1"
)

# This one will list the data collections available in that STAC
collections_query <- stac_source |> 
  collections()

# This runs the collection query and saves the result
coll_list <- get_request(collections_query)

library(tidyverse)
library(purrr)

readable_list <- tibble(
  collection = map_chr(coll_list$collections, "id")
)
```

Now, I need a bounding box. To make that bounding box, I should probably pull in the onX point data of the fields and use the extent of that to create the bounding box for my rstac data query. Then, I will need to convert the points to lines and polygons in order to use the field polygons to clip the sentinel data to just the fields that I want to look at.

```{r LoadData}
#| message: false
#| warning: false

library(terra)

# Loads in all the field data
ReinkeWheelLine <- vect("./data/FieldData/onx-markups-03132026.kml")
LittleReinkePivot <- vect("./data/FieldData/onx-markups-03132026(1).kml")
NReinkeWheelLine <- vect("./data/FieldData/onx-markups-03132026(2).kml")
WilsonPivot <- vect("./data/FieldData/onx-markups-03132026(3).kml")
NorthReinkePivot <- vect("./data/FieldData/onx-markups-03132026(4).kml")
SouthReinkePivot <- vect("./data/FieldData/onx-markups-03132026(5).kml")
UpperPivot <- vect("./data/FieldData/onx-markups-03132026(6).kml")
ClarePivot <- vect("./data/FieldData/onx-markups-03132026(7).kml")
LittleValleyPivot <- vect("./data/FieldData/onx-markups-03132026(8).kml")
SnubPivot <- vect("./data/FieldData/onx-markups-03132026(9).kml")
TurkeyBarnPivot <- vect("./data/FieldData/onx-markups-03132026(10).kml")
BigValleyPivot <- vect("./data/FieldData/onx-markups-03132026(11).kml")
NewValleyPivot <- vect("./data/FieldData/onx-markups-03132026(12).kml")
HousePivot <- vect("./data/FieldData/onx-markups-03132026(13).kml")

# SHOW ME THE MONEY
plot(c(ReinkeWheelLine, LittleReinkePivot, NReinkeWheelLine, WilsonPivot, NorthReinkePivot, SouthReinkePivot, UpperPivot, ClarePivot, LittleValleyPivot, SnubPivot, TurkeyBarnPivot, BigValleyPivot, NewValleyPivot, HousePivot))
```

Well, that didn't quite work like I'd hoped, but it did work well enough to see that there are some problems with this data. Namely, the polygons aren't all clean and round like they should be. Not an immediate problem, but I may need to fix them with some satellite-based geo-referencing before this is over.

```{r CombineData}

# Combines the fields into one spatVector
combined <- rbind(ReinkeWheelLine, LittleReinkePivot, NReinkeWheelLine, WilsonPivot, NorthReinkePivot, SouthReinkePivot, UpperPivot, ClarePivot, LittleValleyPivot, SnubPivot, TurkeyBarnPivot, BigValleyPivot, NewValleyPivot, HousePivot)

# Gets the extent of that spatVector
FieldExt <- ext(combined)

# SHOW ME THE MONEY
plot(combined)
text(combined, 'Name', cex = 0.5 )
```

Okay, well, that's purdy. Got the extent I'm working with. Should be able to go back to `rstac` now and pull some sentinel-1 data.

```{r PullRSTAC}

# Structures the STAC query
stac_query <- stac_search(
  q = stac_source,
  collection = "sentinel-1-rtc",
  bbox = FieldExt[c(1, 3, 2, 4)],
  datetime = "2025-04-01/2025-04-05"
)

# Runs the STAC query
stac_list <- get_request(stac_query)

# Signs the STAC query so Planetary Computer will let me download items
signed_stac_query <- items_sign(
  stac_list,
  sign_planetary_computer()
)

# Downloads a rendered preview
assets_download(signed_stac_query, c("vh", "vv"), 
                output_dir = "./data/RadarData/", overwrite = T)
```

Before we go on, it's worth taking a moment to pause and think about data structures and goals here. I want to download just the data that covers my fields, so I want to use the GDAL system described on the [STAC website](https://stacspec.org/en/tutorials/1-download-data-using-r/) tutorial for using `rstac`. We also need to do some reading about the naming conventions for sentinel data (S1A vs. S1C), and what assets we want (VH, VV). So let's do that and regroup in five.

Okay, let's have a discussion of the strengths and limits of Sentinel-based soil moisture analysis. It is definitely possible to do with a meaningful degree of accuracy, but it requires a few things. First, it requires LAI (Leaf Area Index) control points, which is most accurately determined by doing some insane vegetation cutting and measuring. That's a lot, so I am hoping that since the fields are hay fields with homogeneous grass cover we can get away with estimating LAI from remote sensing data.

Second, there needs to be some input for soil moisture to make this data meaningful. This basically just means soil moisture sensors, which are relatively cheap and easy to install, particularly in hay fields that aren't being tilled every year.

#### (Sound of Brakes Screeching to a Halt)

Scratch all that. Soil moisture may not be possible. Let's just play with the radar data.

```{r RadExp}

radRast <- rast("./data/RadarData/sentinel1-grd-rtc/GRD/2025/4/1/IW/DV/S1C_IW_GRDH_1SDV_20250401T011917_20250401T011942_001693_002D8A_88A4/measurement/iw-vh.rtc.tiff") |> 
  project("epsg:4326")

plot(radRast)
polys(combined)

radRastCrop <- crop(radRast, combined)

plot(radRastCrop, col = map.pal("haxby"))
polys(combined)

bigValleyCrop <- crop(radRastCrop, BigValleyPivot, mask = T)
plot(bigValleyCrop)

snubCrop <- crop(radRastCrop, SnubPivot, mask = T)
plot(snubCrop)
```
