---
title: "Sentinel_Soil_Moisture"
format: html
editor: visual
---

## Sourcing Data

Based on my preliminary readings, it seems like the package `rstac` is going to be my best source for pre-processed and analysis-ready data.

```{r}
#| message: false

library(rstac)

# Sets the STAC to Microsoft's Planetary Computer
stac_source <- stac(
  "https://planetarycomputer.microsoft.com/api/stac/v1"
)

# This one will list the data collections available in that STAC
collections_query <- stac_source |> 
  collections()

# This runs the collection query and saves the result
coll_list <- get_request(collections_query)

library(tidyverse)
library(purrr)

readable_list <- tibble(
  collection = map_chr(coll_list$collections, "id")
)
```

Now, I need a bounding box. To make that bounding box, I should probably pull in the onX point data of the fields and use the extent of that to create the bounding box for my rstac data query. Then, I will need to convert the points to lines and polygons in order to use the field polygons to clip the sentinel data to just the fields that I want to look at.

```{r}
#| message: false
#| warning: false

library(terra)

ReinkeWheelLine <- vect("./data/FieldData/onx-markups-03132026.kml")
LittleReinkePivot <- vect("./data/FieldData/onx-markups-03132026(1).kml")
NReinkeWheelLine <- vect("./data/FieldData/onx-markups-03132026(2).kml")
WilsonPivot <- vect("./data/FieldData/onx-markups-03132026(3).kml")
NorthReinkePivot <- vect("./data/FieldData/onx-markups-03132026(4).kml")
SouthReinkePivot <- vect("./data/FieldData/onx-markups-03132026(5).kml")
UpperPivot <- vect("./data/FieldData/onx-markups-03132026(6).kml")
ClarePivot <- vect("./data/FieldData/onx-markups-03132026(7).kml")
LittleValleyPivot <- vect("./data/FieldData/onx-markups-03132026(8).kml")
SnubPivot <- vect("./data/FieldData/onx-markups-03132026(9).kml")
TurkeyBarnPivot <- vect("./data/FieldData/onx-markups-03132026(10).kml")
BigValleyPivot <- vect("./data/FieldData/onx-markups-03132026(11).kml")
NewValleyPivot <- vect("./data/FieldData/onx-markups-03132026(12).kml")
HousePivot <- vect("./data/FieldData/onx-markups-03132026(13).kml")

plot(c(ReinkeWheelLine, LittleReinkePivot, NReinkeWheelLine, WilsonPivot, NorthReinkePivot, SouthReinkePivot, UpperPivot, ClarePivot, LittleValleyPivot, SnubPivot, TurkeyBarnPivot, BigValleyPivot, NewValleyPivot, HousePivot))
```

Well, that didn't quite work like I'd hoped, but it did work well enough to see that there are some problems with this data. Namely, the polygons aren't all clean and round like they should be. Not an immediate problem, but I may need to fix them with some satellite-based geo-referencing before this is over.

```{r}

combined <- rbind(ReinkeWheelLine, LittleReinkePivot, NReinkeWheelLine, WilsonPivot, NorthReinkePivot, SouthReinkePivot, UpperPivot, ClarePivot, LittleValleyPivot, SnubPivot, TurkeyBarnPivot, BigValleyPivot, NewValleyPivot, HousePivot)

FieldExt <- ext(combined)

plot(combined)
text(combined, 'Name', cex = 0.5 )
```

Okay, well, that's purdy. Got the extent I'm working with. Should be able to go back to `rstac` now and pull some sentinel-1 data.
